Epilepsy example

### 2022-11-08 src/epil/epil.Rmd
beta should be in random, then aghq and TMB match R-INLA method = "gaussian". Almost finished with this case-study

```{r results='hide'}
dat <- list(N = N, J = J, K = K, X = X, y = y, E = make_epsilon_matrix(N, J))

fit1 <- stan(
  "epil.stan",
  data = dat,
  chains = 4,
  warmup = 100,
  iter = 1000,
  control = list(adapt_delta = 0.95)
)
```

@@

We integrate `random` out, which here corresponds to `c("epsilon", "nu")`, with a Laplace approximation:

```{r}
obj <- MakeADFun(
  data = dat,
  parameters = param,
  random = c("epsilon", "nu"),
  DLL = "epil",
  silent = TRUE
)
```

@@

Using just a few of them, we can see that there is a difference, but that difference is a constant so doesn't matter:

@@

```{r results='hide'}
fit8 <- aghq::marginal_laplace_tmb(
  obj,
  k = 3,
  startingvalue = c(param$beta, param$l_tau_epsilon, param$l_tau_nu)
)

summary8 <- summary(fit8)
```

@@

Let's just starting by trying to get the Laplace marginals for $\nu$:

```{r}
compile("epil_nu_index.cpp")
dyn.load(dynlib("epil_nu_index"))
```

@@

Start with `i = 1` for the first marginal:

dat$i <- i

@@

#' Prepare the initialisation parameters
param_fixed_theta[["nu_i"]] <- param_fixed_theta$nu[i]
param_fixed_theta[["nu_minus_i"]] <- param_fixed_theta$nu[-i]
param_fixed_theta[["nu"]] <- NULL

@@

obj_fixed_theta <- MakeADFun(
  data = dat,
  parameters = param_fixed_theta,
  random = c("nu_minus_i", "epsilon"),
  DLL = "epil_nu_index",
  silent = TRUE,
  map = map_fixed_theta
)

@@

ggplot(pdf_and_cdf, aes(x = theta, y = pdf)) +
  geom_line() +
  theme_minimal() +
  labs(x = "nu[1]", y = "Posterior")
```

@@

We're going to use a data structure to pass to `epil_index`.
A `toggle` value of `0` corresponds to `nu` and a toggle value of `1` corresponds to `epsilon`:

```{r}
index_controller <- data.frame(
  "toggle" = c(rep(0, dat$N * dat$J),  rep(1, dat$N)),
  "i" = c(1:(dat$N * dat$J), 1:dat$N)
)
```

@@

Start with `i = 1` for the first marginal:

@@

#' Prepare the initialisation parameters if it's nu that we're toggling
if(dat$toggle == 0) {
  param_fixed_theta[["nu_i"]] <- param_fixed_theta$nu[i]
  param_fixed_theta[["nu_minus_i"]] <- param_fixed_theta$nu[-i]
  # param_fixed_theta$nu <- NULL
}

#' Or if it's epsilon that we're toggling
if(dat$toggle == 1) {
  param_fixed_theta[["epsilon_i"]] <- param_fixed_theta$epsilon[i]
  param_fixed_theta[["epsilon_minus_i"]] <- param_fixed_theta$epsilon[-i]
  # param_fixed_theta$epsilon <- NULL
}

get_random <- function(dat) {
  if(dat$toggle == 0) return(c("nu_minus_i", "epsilon"))
  if(dat$toggle == 1) return(c("nu", "epsilon_minus_i"))
}

@@

# quad <- aghq::aghq(
#   ff = obj_fixed_theta,
#   k = 3,
#   startingvalue = 0,
#   control = aghq::default_control_tmb()
# )
#  
# pdf_and_cdf <- aghq::compute_pdf_and_cdf(quad, finegrid = fine_grid)[[1]]
# 
# ggplot(pdf_and_cdf, aes(x = theta, y = pdf)) +
#   geom_line() +
#   theme_minimal() +
#   labs(x = "nu[1]", y = "Posterior")
```

@@

```{r}
stan1 <- as.vector(t(summary(fit1)$summary[1:6, c(1, 3)]))
inla2 <- as.vector(t(fit2$summary.fixed[1:6, 1:2]))
inla3 <- as.vector(t(fit3$summary.fixed[1:6, 1:2]))
inla4 <- as.vector(t(fit4$summary.fixed[1:6, 1:2]))
tmb5 <- as.vector(t(data.frame(sd_out$par.fixed[1:6], sqrt(diag(sd_out$cov.fixed)[1:6]))))
glmmtmb6 <- as.vector(t(summary(fit6)$coefficients$cond[, c("Estimate", "Std. Error")]))
tmbstan7 <- as.vector(t(summary(fit7)$summary[1:6, c(1, 3)]))
aghq8 <- as.vector(t(summary8$summarytable[1:6, c(1, 4)]))

@@

  ggplot(marginals, aes(x = x, y = y, group = method, col = method)) +
    geom_line() +
    scale_color_manual(values = cbpalette)
    theme_minimal()
}

@@

>>>>

```{r}
dat <- list(N = N, J = J, K = K, X = X, y = y, E = make_epsilon_matrix(N, J))

fit1 <- stan(
  "epil.stan",
  data = dat,
  chains = 4,
  warmup = 100,
  iter = 1000,
  control = list(adapt_delta = 0.95),
  refresh = 0
)
```

@@

We integrate `random` out, which here corresponds to `c("beta", "epsilon", "nu")`, with a Laplace approximation:

```{r}
obj <- MakeADFun(
  data = dat,
  parameters = param,
  random = c("beta", "epsilon", "nu"),
  DLL = "epil",
  silent = TRUE
)
```

@@

Using just a few of them, we can see that the difference between the `TMB` and Stan objectives is a constant:

@@

```{r results='hide'}
fit8 <- aghq::marginal_laplace_tmb(
  obj,
  k = 3,
  startingvalue = c(param$l_tau_epsilon, param$l_tau_nu)
)

fit8_samples <- aghq::sample_marginal(fit8, M = 1000)$samps %>%
  t() %>%
  as.data.frame() %>%
  inf.utils::replace_duplicate_colnames()
```

@@

Let's just starting by trying to get the Laplace marginals for $\beta$:

```{r}
compile("epil_beta_index.cpp")
dyn.load(dynlib("epil_beta_index"))
```

@@

Create a function to get the Laplace marginal for the $i$th element of $\beta$:

```{r}
xi_laplace_marginal_beta <- function(i, opt_theta, finegrid) {
  dat$i <- i

@@

  #' Prepare the initialisation parameters
  param_fixed_theta[["beta_i"]] <- param_fixed_theta$beta[i]
  param_fixed_theta[["beta_minus_i"]] <- param_fixed_theta$beta[-i]
  param_fixed_theta[["beta"]] <- NULL

@@

  obj_fixed_theta <- MakeADFun(
    data = dat,
    parameters = param_fixed_theta,
    random = c("beta_minus_i", "nu", "epsilon"),
    DLL = "epil_beta_index",
    silent = TRUE,
    map = map_fixed_theta
  )

@@

  return(pdf_and_cdf)
}
```

Now we can compute the Laplace marginal for all `r dat$K` indices of $\beta$:

```{r}
fine_grid <- seq(-5, 5, length.out = 1000)

laplace_beta <- lapply(1:dat$K, xi_laplace_marginal_beta, opt_theta = opt_theta, finegrid = fine_grid) %>%
  bind_rows(.id = "index") %>%
  mutate(index = as.numeric(index))
```

Here are what the marginals look like:

```{r}
laplace_beta %>%
  ggplot(aes(x = theta, y = pdf)) +
  geom_line() +
  facet_wrap(~index) +
  theme_minimal() +
  labs(x = "beta", y = "Posterior")
```

@@

We're going to use a data structure to pass to `epil_index`.
A `toggle` value of `0` corresponds to `beta`, a toggle value of `1` corresponds to `nu`, and a toggle value of `2` corresponds to `epsilon`:

```{r}
index_controller <- data.frame(
  "toggle" = c(rep(0, dat$K), rep(1, dat$N * dat$J),  rep(2, dat$N)),
  "i" = c(1:dat$K, 1:(dat$N * dat$J), 1:dat$N)
)
```

@@

Start with `i = 1` for the first marginal of $\beta$:

@@

#' Prepare the initialisation parameters if it's beta that we're toggling
#' We fix the beta parameter to zero, as we're using beta_i and beta_minus_i
if(dat$toggle == 0) {
  param_fixed_theta[["beta_i"]] <- param_fixed_theta$beta[i]
  param_fixed_theta[["beta_minus_i"]] <- param_fixed_theta$beta[-i]
  map_fixed_theta[["beta"]] <- rep(factor(NA), length(param[["beta"]]))
}

#' Or if it's epsilon that we're toggling
#' We fix the epsilon parameter to zero, as we're using epsilon_i and epsilon_minus_i
if(dat$toggle == 1) {
  param_fixed_theta[["nu_i"]] <- param_fixed_theta$nu[i]
  param_fixed_theta[["nu_minus_i"]] <- param_fixed_theta$nu[-i]
  map_fixed_theta[["nu"]] <- rep(factor(NA), length(param[["nu"]]))
}

#' Or if it's epsilon that we're toggling
#' We fix the epsilon parameter to zero, as we're using epsilon_i and epsilon_minus_i
if(dat$toggle == 2) {
  param_fixed_theta[["epsilon_i"]] <- param_fixed_theta$epsilon[i]
  param_fixed_theta[["epsilon_minus_i"]] <- param_fixed_theta$epsilon[-i]
  map_fixed_theta[["epsilon"]] <- rep(factor(NA), length(param[["epsilon"]]))
}

get_random <- function(dat) {
  if(dat$toggle == 0) return(c("beta_minus_i", "nu", "epsilon"))
  if(dat$toggle == 1) return(c("beta", "nu_minus_i", "epsilon"))
  if(dat$toggle == 2) return(c("beta", "nu", "epsilon_minus_i"))
}

@@

ggplot(pdf_and_cdf, aes(x = theta, y = pdf)) +
  geom_line() +
  theme_minimal() +
  labs(x = "beta[1]", y = "Posterior")
```

@@

Let's look at the mean and standard deviation of each of the $\beta$ parameters:

```{r}
stan1 <- as.vector(t(summary(fit1)$summary[1:6, c(1, 3)]))
inla2 <- as.vector(t(fit2$summary.fixed[1:6, 1:2]))
inla3 <- as.vector(t(fit3$summary.fixed[1:6, 1:2]))
inla4 <- as.vector(t(fit4$summary.fixed[1:6, 1:2]))
tmb5 <- as.vector(t(data.frame(sd_out$par.random[1:6], sqrt(sd_out$diag.cov.random[1:6]))))
glmmtmb6 <- as.vector(t(summary(fit6)$coefficients$cond[, c("Estimate", "Std. Error")]))
tmbstan7 <- as.vector(t(summary(fit7)$summary[1:6, c(1, 3)]))

fit8_beta_samples <- fit8_samples[, 1:6]
aghq8 <- as.vector(t(data.frame(mean = apply(fit8_beta_samples, 2, mean), sd = apply(fit8_beta_samples, 2, sd))))

@@

  ggplot(marginals, aes(x = x, y = y, group = method, col = method)) +
    geom_line() +
    scale_color_manual(values = cbpalette) +
    theme_minimal()
}

@@

## `epil_beta_index.cpp`

```{cpp, echo=TRUE, eval=FALSE, output.var="ex3", code=readLines('epil_beta_index.cpp')}
```

## `epil_index.cpp`

```{cpp, echo=TRUE, eval=FALSE, output.var="ex4", code=readLines('epil_index.cpp')}
```
